---
title: "Lab5"
output: html_document
author: Patricia Skora, Taylor O'Rourke
---

# Lab 5

## Problem 0

Our github repo is https://github.com/patryciaskora/MA350Lab5.git. I'm working on RStudio on my personal computer and Taylor is working on RStudio Cloud. 

## Problem 1
```{r}
x <- c(2.2, 7.8, -4.4, 0.0, 1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
quantile(x)
```

The first quartile is -4.525 and the third quartile is 4.9. The interquartile range is Q3 - Q1 which would be 4.9 - -4.525 = 9.425.

## Problem 2
```{r}
#' This function calculates the different quartiles of a set of data using R's built in
#' quantiles() function. It will also calculate the interquartile range based on Tukey's 
#' Rule.
#'
#' @param values A vector containing the values to have quartiles calculated.
#'
#' @return A vector with named components of the first, third, and  interquartile range 
#'         calculated for the argument sent.
quartiles <- function(values){
  quants <- quantile(values)
  first <- quants[["25%"]]
  third <- quants[["75%"]]
  quartRange <- third - first
  
  result <- c("first" = first, "third" = third, "range" = quartRange)
  return(result)
}
quartiles(x)
```
As seen from calling the function above with the vector x, we get the same result as what we calculated above.

## Problem 3

According to Tukey's Law, the points that would be identified as outliers is 28.7 by inspection and the formula.

## Problem 4
```{r}
#' A function that uses the function from above (quartiles()) to calculate the respective 
#' quartiles and interquartile range for a vector of data and checks whether the values 
#' calculated fall under an outlier using Tukey's rule.
#'
#' @param values The vector of data to run the quartile test on.
#'
#' @returns A boolean vector called result that has elements of FALSE if the value in the 
#'          argument passed are not outliers according to Tukey's rule and TRUE otherwrise.
tukey.outlier <- function(values){
  show(values)
  result <- vector(length = length(values))
  response <- quartiles(values)
  intRange <- response["range"]
  Q1 <- response["first"]
  Q3 <- response["third"]
  
  for(val in 1:length(values)){
    if(values[val] < Q1-1.5*intRange | values[val] > Q3 + 1.5*intRange){
      result[val] <- TRUE
    } else{
      result[val] <- FALSE
    }
  }
  return(result)
}
tukey.outlier(x)
```
Our function tukey.outlier works as expected, returning the same TRUE value (28.7) that we calculated above.

## Problem 5
```{r}
#' This function is designed to test a previously written function called tukey.outlier.
#'
#' @param fun A function to be tested against.
#'
#' @return A boolean value of either TRUE if the function correctly identifies the outliers 
#'         and FALSE otherwise.
test.tukey.outlier <- function(fun = tukey.outlier){
  x <- c(2.2, 7.8, -4.4, 0.0, 1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
  testBool <- c(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE)
  functiontest <- fun(x)
  testvec <- testBool == functiontest
  
  return(sum(testvec) == 12)
}
test.tukey.outlier(tukey.outlier)
```

## Problem 6
```{r}
negx <- x * -1
```
Hi Patrycia
